Concurrent, Parallel Garbage Collection in Linear Time

### 2016-06-01 finaletd/mainintroduction.tex
intro finished

\paragraph{Concurrent:}
	Concurrent garbage collectors work simultaneously with application. Application will not experience any pause during execution to scan the memory. They
	have negligible to zero pause time. There are handful of collectors that satisfy this property. 
\paragraph{Multi-collector Garbage :}
	Multi-collector garbage collectors are multiple independent garbage collector thread / process that work independently on the heap to identify garabge. These collectors utilize multiple processors when there is lot of compute resources being under utilized. These type of collectors are essential in distributed systems and shared memory systems. In shared memory systems, when multiple processors are under utilized, this pro
\paragraph{Global Synchronization:}
	When multi-collector garbage collectors are used, conventional solutions require global barrier among all the collectors to communicate and share the information to identify garbage objects and delete them. With no global synchronization, throughput of garbage collectors will be high.
\paragraph{Locality based garbage collection:}
	Detecting garbage is a global predicate. Traditionally, popular methods of garbage collection algorithm involves computing garbage as global predicate. This requires scanning entire heap for better evaluating the predicate. If garbage object can be detected locally based on the a small set of object, then the garbage detection process can be parallelized.
\paragraph{Scalable:}
	With increase in the size of memory, conventional global garbage collection techniques will not scale with synchronization and scanning the entire heap(s). A garbage collection algorithm must be scalable to meet the future demands. Locality based garbage collection with no globally synchronized multi-collector mechanism is the way for scalability.
\paragraph{Promptness:}
	With global garbage collection, the cost of scanning entire heap often is very expensive. It comes with the cost that the garbage will not be collected until the scanning process starts. We require garbage collector that are prompt in collecting garbage as the heap(s) scanning might be expensive in huge distributed systems.
\paragraph{Safety:}
	Safety property is the most crucial part of the problem. Every garbage collector must delete only garbage object. While most garbage collector in literature satisfy this property, there are some that cannot satisfy this property. This thesis requires garbage collector to be safety in the multi-collector environment which is challenging given that there is no global synchronization.
\paragraph{Completeness:}
	Completeness helps garbage collector to claim that it collects all garbage objects in the heap. Reference counting garbage collector is well known for incomplete garbage colletion due to their inability to collect cyclic garbage. This property is challenging as the locality based collectors usually use reference counting. Garbage collector with localized decision making guarantee completeness is necessary to make the collector highly adaptable to enhance the performance of the applications.
\section{Contributions}
\label{intro:contr}
\section{Dissertation Organization}
\label{intro:do}

>>>>

\paragraph{Concurrent:}
	Concurrent garbage collectors work simultaneously with application. Application will not experience any pause during execution to scan the memory. They
	have negligible to zero pause time. There are handful of collectors that satisfy this property. 
\paragraph{Multi-collector Garbage :}
	Multi-collector garbage collectors are multiple independent garbage collector thread / process that work independently on the heap to identify garabge. These collectors utilize multiple processors when there is lot of compute resources being under utilized. In shared memory systems, when multiple processors are under utilized, this property helps to utilize processor cycles better. Apart from the processor utilization, throughput of garbage collectors can be increased by multi-collector garbage collectors. These multi-collector garbage collector threads / process communicate among themselves through memory in shared memory systems and through messages in distributed memory sytems.
\paragraph{Global Synchronization:}
	Multi-collector garbage collectors are common among distributed memory systems. 
	When multi-collector garbage collectors are used, conventional solutions require global barrier among all the collectors to communicate and share the information to identify garbage objects and delete them. With no global synchronization, throughput of garbage collectors will be high and garbage collector algorithm scales well with this property.
\paragraph{Locality based garbage collection:}
	Detecting garbage is a global predicate. Traditionally, popular methods of garbage collection algorithm involves computing garbage as global predicate. This requires scanning entire heap for precise evaluation of the predicate. If garbage object can be detected locally based on the a small set of object, then the garbage detection process can save lot of time spent on tracing all objects in the heap.
\paragraph{Scalable:}
	In shared memory system, conventional single threaded garbage collector does not scale well with increase in size of memory and number of processor. A locality based concurrent multi-collector garbage collector is hope for achieving scalability in shared memory systems.
	With increase in the size of distributed global heap memory, and  number of nodes in distributed memory systems, conventional garbage collection techniques will not scale with global synchronization and scanning the entire heap(s). A distributed garbage collection algorithm must be scalable to meet the future demands. Locality based garbage collection with no globally synchronized multi-collector mechanism is the best bet to scale. Related works in each of the collector chapters describe the scalability scenario in both environments. To my knowledge, there is no completely scalable garbage collectors in both environments.
\paragraph{Promptness:}
	With global garbage collection, the cost of scanning entire heap often is very expensive. To reduce high price for global scanning of the heap, garbage collection is initiated when threshold is met. Promptness property helps to keep heap memory free of floating garbage instantly. This property helps to avoid global scanning and also easy availablity of free memory for allocation.
\paragraph{Safety:}
	Safety property is the most crucial part of the garbage collection which is directly related to advantages of using garbage collection. This avoids any dangling pointers and double free bugs. The property guarantees that every garbage collector will delete only garbage object. While most garbage collector in literature satisfy this property, there are some that cannot satisfy this property. This thesis requires garbage collector to be safety in the multi-collector environment which is challenging given that there is no global synchronization. In distributed memory system, the property is very difficult to satisfy as the garbage collector cannot get the global snapshot of heap at any point in time. Locality based algorithm works only by using information obtained from neighbors. Evaulating a global predicate using local information is challenging by all means. This thesis guarantees the safety of collectors designed for both environments.
\paragraph{Completeness:}
	Completeness helps garbage collector to claim that it collects all garbage objects in the heap, thereby no memory leaks in the heap . Reference counting garbage collector is well known for incomplete garbage collection due to their inability to collect cyclic garbage. This property is challenging as the locality based collectors usually use reference counting. Garbage collector with localized decision making guarantee completeness is necessary to make the collector highly adaptable to enhance the performance of the applications. Collectors designed for both environments uses reference counting based technique to solve the problem and guarentees the garbage collectors are complete.
\section{Contributions}
\label{intro:contr}
The thesis main contributions are listed below :
\begin{enumerate}
	\item Novel reference counting based technique is used to collect any garbage including cyclic garbage.	
	\item First known concurrent multi-collector reference counting based garabge collector is introduced in this thesis.
	\item Well known issues in Brownbridge garbage collectors are fixed and complete and safe garbage collector is designed after three successive failures.
	\item Theoretically, our shared memory garbage collector detects garbage traverses affected graph of objects 7 traversals less than the state of the art concurrent reference counted garbage collector.
	\item This thesis presents first known locality based multi-collector garbage collector with no global synchornization in the distributed memory system.
	\item Distributed garbage collector introduces novel weight based approach to convert the graph into directed acylic graph and thereby detecting cycles faster.
\end{enumerate}
\section{Dissertation Organization}
\label{intro:do}
The preliminaries ideas required to understand the thesis is presented in the chapter ~\ref{prelim}. The chapter ~\ref{prelim} introduces the abstract version of garbage collection problem, brief literature review of garbage collectors in general, and in detail, it explains Brownbridge garbage collector and the reasons for its failure. The understanding of Brownbridge is essential for this thesis as our algorithms extend the approach. The chapter ~\ref{shared} contains the brief introduction of shared memory garbage collectors, related works of the particular class of garbage collectors used in shared memory systems. Shared memory single collector and multi-collector algorithms are described in the chapter ~\ref{shared}. Apart from describing the algorithm, the chapter also contains the proofs of safety and completeness with simulated experimental results. The chapter also proves the linearity in number of operations to detect garbage. The chapter ~\ref{distributed} explains the overview of distributed garbage collection, existing algorithms and their issues. Single collector and multi-collector distributed garbage collector algorithms are explained in abstract and introduces a novel technique to prove termination of garbage collection, safety, completeness and time complexity using isolation property. The chapter also includes the pseudocode of the algorithm. The distributed garbage collector is theoretical leap in solving the problem in with scalability and other guarantees mentioned. The chapter ~\ref{conc} captures the overall contributions of the thesis in broad view with possible future works.
